---
title: "NYC Airbnb Analysis"
author: "Mihika Choudhary, Oyindamola Sanni, Ahlya Elmanzalawi"
date: "2026-05-04"
format: pdf
editor: visual

number-sections: true
toc: true
toc-depth: 4
cap-location: top
fig-align: center
fig-pos: "htbp"

code-annotations: below
code-overflow: wrap

pdf-engine: xelatex
mainfont: "Times New Roman"
monofont: "Courier New"

fontsize: 11pt
linestretch: 1.15

latex-tinytex: true

execute:
  echo: false
  warning: false
  message: false
---

# Introduction

After Airbnb launched in 2008, it slowly began taking over the housing market in New York City, while also becoming a popular lodging option for tourists. Because New York City contains a large variety of neighborhoods and boroughs, we wanted to analyze different listing factors such as location, room type, availability, minimum nights, and ratings to determine which factors most strongly tied to Airbnb popularity.

For our STAT 184 final project, we are analyzing NYC Airbnb listing data to better understand patterns in the Airbnb rental market to better understand what factors influence popular boroughs to rent in. Additionally, we want to determine the most popular booking seasons and draw conclusions from a tourist's perspective to ultimately recommend the best boroughs and seasons for a traveler to book an Airbnb in.

Our main research question is: Which factors appear to influence Airbnb popularity in New York City, and how can these patterns help tourists decide which boroughs and seasons are best for booking an Airbnb?

## Exploratory Data Perspective

For this project, we took an exploratory data analysis approach because our main goal is to identify patterns in the NYC Airbnb rental market rather than just creating cause and effect relationships. To execute this approach, we prioritized analyzing trends, comparing groups, and generating possible explanations for the patterns we find. These results help us make informed recommendations about the best boroughs and seasons to book an Airbnb. Our goal throughout this approach is to identify correlation, not prove causation!

## Datasets and Provenance

For our project, we used Airbnb listing data from Inside Airbnb’s New York City data page:\
https://insideairbnb.com/new-york-city/. We used two raw CSV files from this source: the summary listings dataset and the detailed listings dataset. We then cleaned these files and combined them into our final dataset: 'LisitingReviewsInfo.csv'.

### Primary Dataset: Listings.csv

Our primary dataset throughout this project was `listings.csv` from https://insideairbnb.com/new-york-city/. It provided us with a table summarizing Airbnb listings in New York City, with 36,455 cases and 7 useful attributes for our analysis. Each case in this dataset represents one Airbnb listing. The dataset includes attrivutes about where the listing is located, what type of rental it is, how often it has been reviewed, how available it is, and what minimum stay is required to book.

1\. **Boroughs**: We will use boroughs such as Manhattan, Brooklyn, Queens, the Bronx, or Staten Island to compare which boroughs attract the most tourists and have the most Airbnb activity.\
\
2. **Neighborhood**: We originally assumed boroughs and neighborhoods were relatively the same thing, but with more observation we realized specific neighborhoods lie within each borough. The dataset includes Harlem, Midtown, Hell’s Kitchen, Williamsburg, Long Island city, etc. Since tourists may choose where to stay based on other factors such as transportation, activities, and convenience this attribute helps us look even deeper.\
\
3. **Room type**: These attributes will help us decide what type of rentals tourists prefer, such as: apartments, penthouses, or private rooms.\
\
4. **Minimum night requirement**: This tells us the minimum number of nights a guest must book. Although the data within this attribute was initially shocking, we were able to connect it to recent New York City laws.\
\
5. **Number of reviews and Reviews per month**: These attributes help us understand listing activity, a listing with many reviews may be booked more often, although reviews are not a perfect measure of total bookings because not every guest leaves a review.\
\
6. **Availability 365**: This tells us how many days the listing is available over the next year. This was important to observe from a tourist standpoint because an area with many listings may not actually be easy to book if those listings have low availability.

7\. **Last review**: This gives the date of the most recent review, and we use this attribute to create a new variable called \`last_month_booked\`, which helps us examine seasonal patterns in recent listing activity.

### Secondary Dataset: Detailed NY Listings.csv

We eventually decided to incorporate a more detailed data set to go along with our analysis. Our second dataset is `detailed ny listings.csv`, which also comes from Inside Airbnb’s New York City data page. This dataset contains more detailed information about each Airbnb listing, and later helped us draw more correlations throughout our project.

For our project, we mainly used this dataset to get each listing’s overall guest rating, stored as `review_scores_rating`. The main `listings.csv` file already included review activity information, such as `number_of_reviews`, `last_review`, and `reviews_per_month`, but it did not include the actual star ratings from guests.\
\
To gather this information, we joined this rating data with our cleaned listings dataset using the shared `id` column. This gave us one final dataset that included both general listing information and guest rating information which helped strengthen our hypotheses and broaden our analysis.

## Wrangling Process:

To begin the wrangling process, our team loaded the `dplyr` and `tidyverse` packages so we could clean, filter, and organize the data more efficiently. Before making any changes, we first ran a summary of the original `listings.csv` dataset to get a general understanding of the data, including missing values, possible outliers, and the types of variables we were working with.\
\
After reviewing the dataset, we noticed that several columns were not necessary for our analysis. Since our project focuses on Airbnb popularity, availability, reviews, ratings, and booking patterns, we removed columns such as `name`, `host_profile_id`, `host_name`, `latitude`, `longitude`, `price`, `calculated_host_listings_count`, and `license`. This us helped narrow down the dataset to the attributes that were most relevant to our research question.\
\
Next, we cleaned the remaining data by removing rows with missing values in important columns, especially `minimum_nights` and `reviews_per_month`. We also wanted to focus on listings that were realistically available to book, so we filtered out listings where `minimum_nights` was 0 or `availability_365` was 0. Since our analysis focuses on the main Airbnb rental types, we removed shared rooms and hotel rooms to avoid skewing the results with room types that were not central to our project.\
\
We also noticed outliers in the `minimum_nights` column, with some listings requiring stays longer than 365 days. Since those extreme values could heavily skew our results, we filtered them out and kept listings with minimum-night requirements under one year. Many listings still had minimum stays around 30 nights.\
\
The original `listings.csv` dataset already included review activity variables, such as `number_of_reviews`, `last_review`, and `reviews_per_month`, but it did not include the overall rating score for each listing. To add this information, our team used the detailed listings dataset from Inside Airbnb. We wrangled this file by keeping only the two columns we needed: `id` and `review_scores_rating`. Finally, we were able to join the rating data with our cleaned listings dataset using the shared \`id\` column. Listings without a rating were removed so that our final dataset would include usable rating information.\
\
Finally, we transformed the `last_review` column by separating it into year, month, and day. Since our project also looks at popular booking seasons, we kept the month and removed the day and year. We renamed this new variable `last_month_booked` so it would be easier to interpret in our visualizations. After creating the final combined dataset, we ran another summary to check that the data was cleaned and ready for analysis. We also included comments throughout the code so that the wrangling process would be understandable and reusable.

## F.A.I.R and C.A.R.E Principles:

### **FAIR:**

**Findable**: Both our primary and secondary sources were found on the insideairbnb website. This site is the official data website for AirBNB, making this source reputable.

**Accessible**: Both of our sources were free to use. There was nothing that prevented us from downloading these sources.

**Interoperable**: Our data files were both .csv, which is a widely used format. We were able to easily import our data to start wrangling. Our primary data source had been slightly wrangled by data experts, making the data easy to understand through column names, etc.

**Reusable:** Our wrangled data contains clear and concise comments for readability. This way, another person reading the data can understand and recreate our project.

### **CARE:**

**Collective Benefit:** Our conclusions help renters make decisions while renting in NYC. It is informational and helpful to the collective, and can be utilized to their advantage.

**Authority to Control:** Our data does not violate any property rights, nor does it use highly sensitive information. Our primary dataset did not have price included, which can be linked to privacy reasons.

**Responsibility:** This data has no bias towards any outcome. All sources have been cited.

**Ethics:** Our data does not promote any form of harm, nor does it exploit others. No person is harmed, no laws are broken, and our findings are morally sound. This data can be used for future use.

```{r}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(knitr)
library(kableExtra)

listings <- read.csv("listings.csv")
detailed.ny.listings <- read.csv("detailed ny listings.csv")
ListingReviewsInfo <- read.csv("ListingReviewsInfo.csv")
```

## Summary Table by Borough

Before analyzing individual statistics, we first summarized important Airbnb listing characteristics by borough. This table allows us to see the exacrt number of listings, average reviews per month, average review score, average availability, and average minimum night requirement. Examining these exact numbers gave us a general idea regarding each borough based on both popularity and tourist preferences.

```{r}
#| label: tbl-borough-summary
#| tbl-cap: "Summary Statistics by NYC Borough for Airbnb Listings"
#| results: asis

borough_summary <- ListingReviewsInfo |>
  group_by(neighbourhood_group) |>
  summarise(
    `Listings (n)` = n(),
    `Avg Reviews/Month` = round(mean(reviews_per_month, na.rm = TRUE), 2),
    `Avg Review Score` = round(mean(review_scores_rating, na.rm = TRUE), 2),
    `Avg Availability (days)` = round(mean(availability_365, na.rm = TRUE), 0),
    `Avg Min. Nights` = round(mean(minimum_nights, na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  arrange(desc(`Avg Reviews/Month`)) |>
  rename(Borough = neighbourhood_group)

cat("\\begin{center}\n")

borough_summary |>
  kable(
    format = "latex",
    booktabs = TRUE,
    align = "c"
  ) |>
  kable_styling(
    latex_options = c("hold_position", "striped", "scale_down"),
    full_width = FALSE
  )

cat("\\end{center}\n")
```

**Alt Text**: Table showing borough-level Airbnb statistics for Queens, Brooklyn, Manhattan, the Bronx, and Staten Island, including listing count, average reviews per month, average rating, average availability, and average minimum-night stay. **\*Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.\***

## Statistical Analysis:

Table 1 summarizes our chosen attributes for Airbnb listings per borough. Strictly based on numbers, we can see that Manhattan has the most amount of listings at 5240, and Brooklyn comes in second place with 4762 listings. Manhattan and Brooklyn seem to be dominating the Airbnb market which is likely due to the tourist attractions, restaurants, scenery, and popular stores they offer. However, Queens has the highest review activity at 1.07 reviews per month, followed closely by Brooklyn at 1.06 and Manhattan at 1.04. This suggests that Queens listings may be slightly more active, even though Queens has fewer listings than Manhattan and Brooklyn.

### Long Description:

This table compares five Airbnb listing measures across NYC boroughs. Manhattan and Brooklyn stand out for having the largest number of listings, which suggests that tourists have the most Airbnb options in those boroughs. Queens, however, has the highest average reviews per month, meaning its listings show slightly stronger review activity even though it has fewer total listings than Manhattan and Brooklyn. Staten Island has the highest average review score and the highest average availability, but it also has the smallest number of listings, so its results should be interpreted with caution. The Bronx also has high availability compared to the larger boroughs, which may suggest that listings there are easier to find but possibly less in demand. Across all boroughs, the average minimum-night requirement is high, ranging from about 25 nights to about 30 nights, showing that many listings may be less practical for short tourist stays. Overall, the table shows that no single borough is strongest in every category, so tourist recommendations should consider listing count, review activity, ratings, availability, and minimum-night requirements together. **\*Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.\***

## Median Minimum Booking Nights by Borough

After we successfully summarized our borough statistics, we were quite confused seeing the average minimum booking night requirement per listing. We created a visualization to get a better idea on what each listing requires as minimum night booking on average.

```{r}
#| fig-cap: "Figure 1: Median Minimum Booking Nights by Borough"
#| fig-alt: "Bar chart showing median minimum booking nights for Airbnb listings across New York City boroughs. Boroughs are on the x-axis, median minimum booking nights are on the y-axis, and the y-axis ranges from 0 to 50 nights."

minimum_nights_by_borough <- ListingReviewsInfo |>
  group_by(neighbourhood_group) |>
  summarize(
    median_minimum_nights = median(minimum_nights, na.rm = TRUE),
    .groups = "drop"
  )

minimum_nights_by_borough |>
  ggplot(aes(
    x = reorder(neighbourhood_group, -median_minimum_nights),
    y = median_minimum_nights,
    fill = neighbourhood_group
  )) +
  geom_col() +
  scale_y_continuous(limits = c(0, 50)) +
  labs(
    title = "Median Minimum Booking Nights by Borough",
    x = "Borough",
    y = "Median Minimum Booking Nights"
  ) +
  theme_minimal() +
  theme(legend.position = "none")
```

**Alt Text**: Bar chart showing the median minimum booking nights for Airbnb listings across the Bronx, Brooklyn, Manhattan, Queens, and Staten Island. Each borough has the same median minimum stay of 30 nights. **\*Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.\***

## Statistical Analysis:

Observing the bar graph, we can see that median minimum booking requirement is 30 nights across every borough. At first, this month long requirement seemed unusual, especially from a tourist’s perspective, since many visitors usually look for shorter stays. We were later able to connect this to a recent New York City law. New York City adopted Local Law 18 back in September 2023. This law is also known as the "Short-term Rental Registration Law". Local Law 18 placed restrictions on Airbnb hosts providing short term stays in their listings, meaning they would have to register with the Mayor's Office of Special Enforcement if they wanted to offer rental stays under 30 consecutive days. We can assume many hosts may now choose to set their minimum stay at 30 nights to avoid the additional registration process to offer short term stays to their guests. An insight we had throughout this analysis was that short term visits to New York City may not be realistic for travelers through Airbnb, and they may choose to book a hotel instead. Although our data does not intend to prove that this law directly caused the 30 night minimum requirement, it provides us with a strong possible explanation for why this pattern seems to appears across our entire dataset.

### Long Description:

This bar chart compares the median minimum booking-night requirement across all five New York City boroughs. The x-axis lists the Bronx, Brooklyn, Manhattan, Queens, and Staten Island, while the y-axis shows median minimum booking nights from 0 to 50. Each borough has a bar reaching 30 nights, meaning the typical listing in every borough requires a minimum stay of about one month. This is an important pattern because it shows that minimum-night requirements are consistent across boroughs rather than being limited to one specific area. From a tourist standpoint, this may make many Airbnb listings less practical for short trips, since travelers visiting for only a few days or a week would not be able to book listings with a 30-night minimum. Overall, the graph suggests that tourists may need to pay close attention to minimum-night requirements when deciding where to book, because this restriction appears across all boroughs in the dataset. **\*Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.\***

## Listing Popularity by Borough

Our Figure 2 visualization compares listing popularity by borough using average reviews per month. Since our dataset did not include the exact number of bookings, we used reviews per month as a helpful substitution to determine how active Airbnb listings are in each borough.

```{r}
borough_pop <- ListingReviewsInfo |>
  group_by(neighbourhood_group) |>
  summarise(avg_reviews_pm = mean(reviews_per_month, na.rm = TRUE)) |>
  mutate(neighbourhood_group = fct_reorder(neighbourhood_group, avg_reviews_pm))

plot1 <- ggplot(borough_pop, aes(x = neighbourhood_group,
                                 y = avg_reviews_pm,
                                 fill = neighbourhood_group)) +
  geom_col(show.legend = FALSE, width = 0.6) +
  geom_text(aes(label = round(avg_reviews_pm, 2)),
            vjust = -0.5, size = 4, fontface = "bold") +
  scale_fill_manual(values = c(
    "Staten Island" = "#8ecae6",
    "Bronx"         = "#219ebc",
    "Manhattan"     = "#023047",
    "Brooklyn"      = "#ffb703",
    "Queens"        = "#fb8500"
  )) +
  scale_y_continuous(
    limits = c(0.7, 1.2),
    oob    = scales::squish
  ) +
  labs(
    title    = "Listing Popularity by NYC Borough",
    subtitle = "Measured by average number of reviews per month",
    x        = "Borough",
    y        = "Average Reviews per Month",
    caption  = "Source: NYC Airbnb listings dataset"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    plot.title         = element_text(face = "bold", size = 14),
    plot.subtitle      = element_text(color = "gray40", size = 11),
    plot.caption       = element_text(color = "gray50", size = 9),
    panel.grid.major.x = element_blank(),
    panel.grid.minor   = element_blank(),
    axis.text.x        = element_text(size = 11)
  )

plot1
```

**Alt Text**: Bar chart comparing average reviews per month for Airbnb listings across Staten Island, the Bronx, Manhattan, Brooklyn, and Queens. Queens has the highest average reviews per month at 1.07, followed by Brooklyn at 1.06 and Manhattan at 1.04. The Bronx and Staten Island have lower averages at 0.86 and 0.85. ***Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.***

## Statistical Analysis:

Based on the graph, we were able to get a visual idea of listing popularity across New York City. Queens has the highest average review activity at 1.07 reviews per month, with Brooklyn close behind at 1.06 and Manhattan at 1.04. The Bronx and Staten Island are somewhat lower, showing Bronx at 0.86 and Staten Island at 0.85.

This visualization suggests that Queens, Brooklyn, and Manhattan may have more active Airbnb listings compared to the Bronx and Staten Island. However, the top three most reviewed boroughs (Queens, Brooklyn, and Manhattan) contain extremely small differences, so we can't confidently say that this category necessarily shows us the most popular borough. From a tourist’s perspective, these three boroughs seem like the best booking options because their listings are reviewed most often and that could likely provide a sense of reassurance prior to booking. The Bronx and Staten Island may still be good options depending on the traveler's needs, but their lower review activity suggests they may not be as popular or active in the Airbnb market.

### 

### Long Description:

This bar chart shows average Airbnb reviews per month by New York City borough. The x-axis lists Staten Island, the Bronx, Manhattan, Brooklyn, and Queens, while the y-axis shows average reviews per month from 0.7 to 1.2. Queens has the tallest bar at 1.07 reviews per month, with Brooklyn close behind at 1.06 and Manhattan at 1.04. The Bronx and Staten Island have shorter bars, at 0.86 and 0.85 reviews per month. The chart suggests that Queens, Brooklyn, and Manhattan have higher review activity than the Bronx and Staten Island. Since reviews per month are being used as a proxy for listing activity, this graph helps compare which boroughs may have more frequently used Airbnb listings. However, the y-axis starts at 0.7, so the visual difference between boroughs may appear larger than the numerical difference actually is. ***Written with assistance from ChatGPT, GPT-5.5 Thinking, on May 6, 2026.***
